---
title: "Processing Script"
author: "William Norfolk"
date: "10/26/2019"
output: word_document
---


This Rmd file loads, cleans, and organizes data variables collected from the Fall 2019 EHSC 8310 water sampling.

Start by loading all required libraries. 

```{r load libs}

library(readxl)
library(dplyr)
library(tidyverse)
library(forcats)
library(ggthemes)
library(knitr)
library(naniar)
library(visdat)


```

Next load the data. The path in the code chunk below is relative to the raw data file placed in this project. 

```{r load data}
micro_raw_data <- readxl::read_excel("./raw_data.xlsx")
```

Take a look at the data and variables to see what you are working with.

```{r glimpse}
glimpse(micro_raw_data)
```


Lots of variables here, mostly just information that needed to be recorded but is not particularly useful for data analysis so we will remove some variables that are not needed for visualization/statistics for simplicity sake. Also a number of the variables loaded into the data set with strange additions so we will remove them to reduce typos.

```{r}
clean_names<- micro_raw_data %>% rename(sample_id = `Sample ID`,
                                     collection_time = `Collection Time`,
                                     ambient_air_high_c = `Ambient Air Temp High (C)`,
                                     ambient_air_low_c = `Ambient Air Temp Low (C)`,
                                     water_temp = `Water Temp (C)`,
                                     conductivity = `Conductivity (mS/cm)`,
                                     final_ecoli = `Final CFU/100 ml`,
                                     exceeds_epa = `Exceeds EPA STV (410/100mL)`,
                                     salmonella = `Chromagar Confirmed Salmonella`,
                                     date = Date,
                                     ph = pH)

```


```{r}
reduce_vars <- clean_names %>% dplyr::select(date, ambient_air_high_c, ambient_air_low_c, water_temp, conductivity, ph, final_ecoli, exceeds_epa, salmonella )

glimpse(reduce_vars)
```

Now we can check what kind of data and observations we are dealing with.

```{r}
vis_dat(reduce_vars)
```

